---
title: "p8105_hw1_dap2189"
output: github_document
date: "2025-09-13"
---

# Problem 1

This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.

```{r}
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

Note: 
I used below line in the console to see the first 6 rows (making sure I installed it correctly: `head(early_january_weather)`


This dataset looks at weather conditions in early January. This dataset has `r nrow(early_january_weather)` rows, `r ncol(early_january_weather)` columns, and a mean temperature of `r round(mean(early_january_weather$temp, na.rm = TRUE),2)` degrees.

Variables in this dataset include: `r names(early_january_weather)`. A summary of the variables and significant values can be seen below.

```{r, echo=FALSE}
summary(early_january_weather)
```

## Creating a Scatter Plot

Based on the below scatterplot, it appears that temperature increases as we get later into January. As well, there is a similar trend with humidity: later in the month when the temperatures are higher, the humidity index increases. 

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()

ggsave("january_weather_and_humidity.pdf", height = 4, width = 6)
```


# Problem 2

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

```{r}
Problem2_df = tibble(
    random_number = rnorm(10),
    logical_greater_than_0 = random_number > 0,
    character_var = c("My", "name", "is", "Jeff", "and", "I", "a","professor", "at", "Mailman"),
    factor_size = factor(c("Small", "Medium", "Large", "Small", "Medium", "Large","Small", "Medium", "Large","Small"))
)

```

Notes on creating a data frame comprised of:
factor_var = factor(c("male", "male", "female", "female")) --> is there a way to do this without writing out all 10 times?'

## Taking the Mean

Take the mean of each variable in your dataframe. What works and what doesn’t?
```{r, echo=TRUE, warning=FALSE}
mean(pull(Problem2_df,random_number))
mean(pull(Problem2_df,logical_greater_than_0))
mean(pull(Problem2_df,character_var))
mean(pull(Problem2_df,factor_size))
```

We needed to use the pull function to pull the variable out of the dataset. 

The mean of variable, "random_number" is `r mean(pull(Problem2_df,random_number))` and the mean of variable, "logical_greater_than_0" is `r mean(pull(Problem2_df,logical_greater_than_0))`. I cannot get the mean of variables "character_var" and "factor_size" becuase they are non-numeric values. The logical variable can be used in calculations because it gets a value of 1 if random_variable is greater than 0 or 0 if less than 0.

## Forcing Variables

Using `as.numeric()` to force a variable from original form into numeric:

```{r, echo=TRUE, results='hide', warning=FALSE}
as.numeric(Problem2_df$logical_greater_than_0)
as.numeric(Problem2_df$character_var)
as.numeric(Problem2_df$factor_size)
```

With the logical variable, it prints out values of 0 and 1. 
With the character variable, it says "NA". I think this is because it cant convert text to numbers.
With the factor variable, because it creates three categories, perhaps it assigns a value of 1-3 to each string value, and then outputs the number associated with the category. (ex. 3=small, 2=medium, 1=large)
